---
title: "Assignment 3 problem 1"
author: "SRK Yarra"
date: "2023-10-25"
output: html_document
---
```{r}
#Problem1(a)
# Read the data
data <- read.csv("college.csv")

#a) Analyze distribution of Grad.Rate
# Load necessary libraries
library(ggplot2)
# Plot the distribution of Grad.Rate
ggplot(data, aes(x=Grad.Rate)) + geom_histogram(binwidth=5, fill="blue", color="black") +
  labs(title="Distribution of Grad.Rate", x="Graduation Rate", y="Frequency") +
  theme_minimal()

data$Grad.Rate<-log(data$Grad.Rate)

ggplot(data, aes(x=Grad.Rate)) + geom_histogram(binwidth=0.1, fill="blue", color="black") +
  labs(title="Transformed Distribution of Grad.Rate", x="Graduation Rate", y="Frequency") +
  theme_minimal()
```

#Problem1(b)

```{r}
#b) Scatterplots for Grad.Rate vs each independent variable
# Creating scatterplots for each variable
independent_vars <- names(data)[2:(ncol(data)-1)] # Exclude 'school' and 'Grad.Rate'
for (var in independent_vars) {
  print(ggplot(data, aes_string(x=var, y="Grad.Rate")) + geom_point(alpha=0.5) + theme_minimal())
}

# Correlation analysis

correlation_matrix<-cor(data[,-c(1,2)])
correlation_matrix
```

#Problem1(c)
```{r}
#c) Boxplots for graduation rates by university type and elite status

# Boxplot for Grad.Rate by Private/Public University
ggplot(data, aes(x=Private, y=Grad.Rate)) +
  geom_boxplot(fill="blue", color="black") +
  labs(title="Grad.Rate by University Type", x="University Type", y="Graduation Rate") +
  theme_minimal()

# Boxplot for Grad.Rate by Elite/Not Elite Status
ggplot(data, aes(x=factor(Elite10), y=Grad.Rate)) +
  geom_boxplot(fill="green", color="black") +
  labs(title="Grad.Rate by Elite Status", x="Elite Status (1=Elite, 0=Not Elite)", y="Graduation Rate") +
  theme_minimal()
```

#Problem1(d)
```{r}
#d) Fit a full model
full_model <- lm(Grad.Rate ~ . - school, data=data)
summary(full_model)
```

#Problem1(e)
```{r}
#e) Multi-collinearity and VIF statistics
library(car)

vif_values <- vif(full_model)
print(vif_values)
```

#Problem1(f)
```{r}
#f) Variable selection procedures
# Backward Selection
backward_model <- step(full_model, direction="backward")

# Forward Selection
null_model <- lm(Grad.Rate ~ 1, data=data)
forward_model <- step(null_model, scope=list(lower=null_model, upper=full_model), direction="forward")
```

#Problem1(g)
```{r}
#g) Fit a final regression model M1
#final Model based on the backward selection as observed in the output
M1 <- lm(Grad.Rate ~ Private + Accept.pct + Elite10 + F.Undergrad + P.Undergrad +
           Outstate + Room.Board + Personal + PhD + perc.alumni + Expend, data=data)
summary(M1)
```

#Problem1(h)
```{r}
#h) Scatter plot of studentized residuals against predicted values
# Compute studentized residuals
studentized_residuals <- rstudent(M1)

# Scatter plot
ggplot(data, aes(x=predict(M1), y=studentized_residuals)) +
  geom_point(alpha=0.5) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(title="Studentized Residuals vs. Predicted Values", x="Predicted Values", y="Studentized Residuals") +
  theme_minimal()
```
#Problem1(i)
```{r}
#i) Normal probability plot of residuals
qqnorm(studentized_residuals)
qqline(studentized_residuals)
```
#Problem1(j)
```{r}
#j) Outliers or Influential Points
# Cook's distance to detect influential observations
cook_d <- cooks.distance(M1)

a<-influence.measures(M1)
summary(a)
plot(rstudent(M1)~hatvalues(M1))
abline(a=3, b=0, col= 'red')
abline(a=-3, b=0,col='red')

# Flag observations with Cook's distance > 4/n
influential_points <- as.numeric(names(cook_d)[(cook_d > 4/length(cook_d))])
print(influential_points)

# For outliers, you can inspect large studentized residuals
outliers <- which(abs(studentized_residuals) > 2) # Adjust the threshold as necessary
print(outliers)
```
#Problem1(k)
```{r}
#k) adjusted r2 value
r_squared <- summary(M1)$r.squared
print(r_squared)
```
#problem 1(L)
Being an Elite10 university significantly affects the relationship between graduation rates and the predictors. Specifically:

The negative effect of acceptance percentage on graduation rate is stronger for Elite10 universities.

The positive effect of out-of-state tuition on graduation rate is weaker for Elite10 universities.

The relationship between the percentage of alumni who donate and graduation rate doesn't change significantly for Elite10 universities.

The negative relationship between expenditure per student and graduation rate reverses for Elite10 universities.

This means that the factors influencing graduation rates differ somewhat for Elite10 universities compared to other institutions.





